# Run Performance and Read Depth

```{r}
counts_by_sample <- read.csv(file.path(ROOT, "reporting/counts_by_sample.csv"), header = TRUE, stringsAsFactors = FALSE)
counts_by_sample <- subset(counts_by_sample, Sample != "unassigned")
counts_by_run <- read.csv(file.path(ROOT, "reporting/counts_by_run.csv"), header = TRUE, stringsAsFactors = FALSE)
counts_by_amplicon <- read.csv(file.path(ROOT, "reporting/counts_amplicon_summary.csv"), header = TRUE, stringsAsFactors = FALSE)
counts_assembly <- read.csv(file.path(ROOT, "reporting/counts_assembly_summary.csv"), header = TRUE, stringsAsFactors = FALSE)
counts_qual <- read.csv(file.path(ROOT, "reporting/counts_presto_qual_summary.csv"), header = TRUE, stringsAsFactors = FALSE)
```

```{r}
counts_presto <- subset(counts_by_amplicon, select = -Ratio)
handle <- with(counts_presto, paste(Timepoint, Specimen, Chain, ChainType))
if (! (length(handle) == length(unique(handle)))) {
  stop("Non-unique IDs for specimen handling")
}

idx <- match(handle, with(counts_assembly, paste(Timepoint, Specimen, Chain, ChainType)))
counts_presto$SeqsAssembled <- counts_assembly$Seqs[idx]
idx <- match(handle, with(counts_qual, paste(Timepoint, Specimen, Chain, ChainType)))
counts_presto$SeqsQualFilt <- counts_qual$Seqs[idx]

counts_presto <- counts_presto[, c(colnames(counts_presto)["CellCount" != colnames(counts_presto)], "CellCount")]
counts_presto$FinalRatio <- with(counts_presto, round(SeqsQualFilt/CellCount, 8))
```

```{r}
drawtab <- function(df) {
  kableExtra::kable_styling(
    knitr::kable(df, digits = 3, row.names = FALSE, booktabs = TRUE))
}
```

## By Sequencing Run

We aim for a 50/50 ratio of library to PhiX. Assuming the unassigned reads are
all PhiX, we expect to see close to 1.0 for the below ratios.

```{r}
knitr::kable(counts_by_run, digits = 3, row.names = FALSE)
```

## By Sequencing Sample

The libraries are prepared with a goal of roughly ten reads for every one cell,
so we expect to see 10 or more for the below ratios.

```{r}
knitr::kable(counts_by_sample, digits = 3, row.names = FALSE)
```

## By Sequencing Sample - Adapter and Quality Trimming

```{r}
qualgrids_r1 <- list()
qualgrids_r2 <- list()
for (idx in 1:nrow(counts_by_sample)) {
  prefix <- file.path(
    ROOT,
    "qc",
    counts_by_sample$Run[idx],
    paste0("qualtrim.", counts_by_sample$Sample[idx]))
  qualgrids_r1 <- c(qualgrids_r1, list(load_qualgrid(paste0(prefix, ".R1.csv"))))
  qualgrids_r2 <- c(qualgrids_r2, list(load_qualgrid(paste0(prefix, ".R2.csv"))))
}
names(qualgrids_r1) <- counts_by_sample$Sample
names(qualgrids_r2) <- counts_by_sample$Sample
```

```{r fig.width = 3.7, fig.height = 3, dev="png", dpi=150}
for (sample in counts_by_sample$Sample) {
  if (sum(qualgrids_r1[[sample]]) > 100) {
  plot_qualgrid(qualgrids_r1[[sample]], main=paste(sample, "R1"),
                legend = FALSE, show_rownames = FALSE)
  plot_qualgrid(qualgrids_r2[[sample]], main=paste(sample, "R2"))
  cat("\n\n")
  } else {
    cat(paste("\n\nToo few reads for", sample, "\n\n"))
  }
}
```
